# Setup

To get started, ensure you have a recent verison or R and RStudio installed. 

## Software  

### Install R  

To install R head to [https://cran.rstudio.com/](https://cran.rstudio.com/). 

### Install RStudio  

Next, install RStudio Desktop IDE at [https://posit.co/download/rstudio-desktop/](https://posit.co/download/rstudio-desktop/)

### Packages  

To install the required packages, run:  
```{r}
#| eval: false
install.packages("tidyverse")
install.packages("mgcv")
install.packages("pins")
install.packages("vetiver")
install.packages("plumber")
install.packages("here")
install.packages("roxygen2")
install.packages("testthat")
```

You should be able to now run the following commands: 

```{r}
#| eval: false
library(tidyverse)
library(mgcv)
library(pins)
library(vetiver)
library(plumber)
library(here)
library(roxygen2)
library(testthat)
```

## Code  

### Download from Github  

Materials for the case study can be cloned by using the following command from the `usethis` package:

Alternatively visit: [https://github.com/deanmarchiori/beachwatch](https://github.com/deanmarchiori/beachwatch)

```{r}
#| eval: false
usethis::create_from_github("deanmarchiori/beachwatch", fork = FALSE)
```

